# SAPA project website

## Introduction
The [SAPA Project](https://sapa-project.org/) hosts surveys that explore different dimensions
of personality.

### Infrastructure
The SAPA Project is hosted on [Google Cloud](https://cloud.google.com/). There is one active compute engine virtual machine (VM) instance, and one backup or staging instance. The VM instances run an [Apache/2.4.25 web server](http://httpd.apache.org/docs/2.4/) on Debian 9 (stretch) Linux based operating system that hosts the actual SAPA Project website. Each time a user completes a survey, the survey results are stored in a MySQL database, which is running on a Google Cloud SQL instance running [MySQL 5.7](https://dev.mysql.com/doc/refman/5.7/en/).

#### Admin console
Admin console for the computing infrastructure is available at [https://console.cloud.google.com](https://console.cloud.google.com).

#### How to add new administrators / owners
Add a new user from the [IAM & Admin](https://console.cloud.google.com/iam-admin/iam?project=silken-alloy-248920) admin console page.

### How to connect to database

*First time users:*

1. Download and install the [Google Cloud SDK](https://cloud.google.com/sdk/docs/downloads-interactive)
2. Download and install the [Google Cloud SQL Proxy](https://cloud.google.com/sql/docs/mysql/connect-admin-proxy#install)
3. Download and install a MySQL client. On macOS, use homebrew to install a command line client via `brew install mysql`, or you can install the [MySQL Workbench GUI client](https://dev.mysql.com/downloads/workbench/)



* In a terminal session, authenticate to Google services
```{bash, eval=FALSE}
gcloud auth login
```
A web browser will open and direct you to login to your Google account.

* Start the `cloud_sql_proxy`:
```{bash, eval=FALSE}
cloud_sql_proxy -instances=silken-alloy-248920:us-west1:woodworth-pi-improvement-project-1920=tcp:3306
```

If the command succeeds, you'll see the following message returned to your shell:
```{bash, eval=FALSE}
Ready for new connections
```
The Cloud SQL Proxy allows you to make connections to the MySQL database running in Google Cloud as if the database was running locally. 

* In a new terminal window, verify the connection:
```{bash, eval=FALSE}
mysql -h 127.0.0.1 -u SAPAreader -p
```
Enter the password when prompted. If the connection is successful, then you'll see the mysql prompt:
```{bash, eval=FALSE}
mysql> 
```

### Connecting from R
Following the instructions here: https://cloud.google.com/blog/products/gcp/google-cloud-platform-for-data-scientists-using-r-with-google-cloud-sql-for-mysql

It is helpful to create a function that creates the connection to the MySQL database:
```{r eval=FALSE}
library(RMySQL)

getSqlConnection <- function(){
  con <-
    dbConnect(
      RMySQL::MySQL(),
      username = 'SAPAreader',
      password = 'password',
      host = '127.0.0.1',
      dbname = 'SAPAactive'
    )
  return(con)
}
```

Then you can query any table in the `SAPAactive` database by following this example, which queries 10 results from the  `TAIE_responses_111119` table:
```{r eval=FALSE}
conn <- getSqlConnection()
res <- dbSendQuery(conn, "select * from TAIE_responses_111119 limit 10")
data <- dbFetch(res)
print(data)
dbDisconnect(conn)
```
